sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-03-16 inst/documents/rotations_notes.Rmd
sensi analysis

*1/2/2018*

I want to create a figure that shows effect of untreated area, refugia
Can I do this in ggplot2, or do it in powerpoint ?
see : resistance-dispersal-dilution-figure.pptx

@@

re-read this 
Sternberg & Thomas 2017
Insights from agriculture for the management of insecticide resistance in disease vectors
Because of the situation-specific nature of insecticide resistance management plans, public health may have an advantage over agriculture. Agricultural pest management is concerned with hundreds
of species, and in many cases, detailed data are not available for each pest species (and potential nontarget organisms). Compared to agriculture, public health is concerned with a small number of insect species. This should enable a strongly data-driven approach to selecting resistance management strategies, but in reality, very little is currently known.

@@

Download papers on modelling oxitech
Insectcide resistance
https://doi.org/10.1093/jee/100.5.1642
https://doi.org/10.1603/029.102.0233

@@

send Rowland rotation paper to Ian.

@@

~ c 15 generations per year for mosquitos, 25 day generation time.
~ 3 years = 45 generations
~ https://www.cdc.gov/malaria/about/biology/mosquitoes/
~

>>>>

*1/2/2018*
I want to create a figure that shows effect of untreated area, refugia
Can I do this in ggplot2, or do it in powerpoint ?
see : resistance-dispersal-dilution-figure.pptx

@@

~ c 15 generations per year for mosquitos, 25 day generation time.
~ 3 years = 45 generations
~ https://www.cdc.gov/malaria/about/biology/mosquitoes/

*14/3/2018*
re-red this 
Sternberg & Thomas 2017
Insights from agriculture for the management of insecticide resistance in disease vectors
Because of the situation-specific nature of insecticide resistance management plans, public health may have an advantage over agriculture. Agricultural pest management is concerned with hundreds
of species, and in many cases, detailed data are not available for each pest species (and potential nontarget organisms). Compared to agriculture, public health is concerned with a small number of insect species. This should enable a strongly data-driven approach to selecting resistance management strategies, but in reality, very little is currently known.

@@

*15/3/2018* back from Peru and Geneva
~ get trial run of sensitivity analysis going
~ sensi_an_rotations1.Rmd

~ now have 500 runs working, max gens up to 400
~ seems that generations are identical for sequence and rotations
~ results saved in df_res_all_rot.rda

Setting coverage to <1 (0.5) and migration to 0.01 does result in rotations lasting longer than sequences in some scenarios. Particularly shorter ones. Greatest slope is when rot_interval set to 5, one scenario gives 300 for sequence and 350 for rotation.

So maybe difference is created by the changeover, and more frequent changeovers create greater difference.

Also when coverage set to 1 again and cost > 0 there was 1 scenario with a very short rot-interval where the rotation lasts fair bit longer than the sequence.

## scenario 79 expo_hi 0.79 eff 0.72 rot_interval 0
## tot gens deployed under freq 0.5 = 199

## scenario 79 expo_hi 0.79 eff 0.72 rot_interval 6
## tot gens deployed under freq 0.5 = 251

run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)

It could perhaps be just a rounding thing, because the number of generations is an integer there will always be some overshooting, when you have more switches there is the potential for more overshooting and the runs will last for longer.

~ checking on coverage=0, currently resistance frequencies increase when they shouldn't do 
run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=0, add_gens_under50=TRUE)

in run_rot()

coverage is passed to :
set_start_freqs() but only to determine whether the

RAF 3rd dimension site is set to : c('intervention','refugia') or just 'intervention'

actually if coverage=0 perhaps dimension should just be set to refuge.

So maybe it doesn't point to an error that resistance rises when coverage=0 just a property of the model structure.

~ in run_rot finish code condensing around :
      raf_m <- RAF[insecticide, 'm', 'intervention',gen-1]
      raf_f <- RAF[insecticide, 'f', 'intervention',gen-1]

~ sensi_an_rotations1.Rmd

~ Maybe don't allow coverage to be set to 0 ? or put in some kind of warning.

~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)

~ can I create an easy function to compare two scenarios (seq & rotate) with otherwise same params ?

Download papers on modelling oxitech
Insecticide resistance
https://doi.org/10.1093/jee/100.5.1642
https://doi.org/10.1603/029.102.0233

@@

@@

### 2018-03-16 inst/documents/sensi_an_rotations1.Rmd
sensi analysis

n_scenarios <- 5

@@

for(i in 1:n_scenarios)
{
  # expo_hi 0.1 – 0.9
  # eff 0.3 - 1
  linputs <- list()

  linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
  linputs$eff <-      runif(1, min=0.3, max=1)

  # run one scenario
  dfres <- do.call(run_rot, linputs)

  # summarise results per scenario
  res <- dfres %>%
  dplyr::filter(active_or_refuge=='active') %>%
  group_by(resist_gene) %>%
  summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
  summarise(mean_gens_under50 = mean(gens_under50)) %>%
  unlist()

  # add results onto inputs list
  linputs <- c(linputs, res)

  # change list to dataframe
  df_in_out <- as.data.frame(linputs)

@@

}

@@

<!-- eval=FALSE to stop slow code being executed after 1st time --> 
```{r, eval=FALSE}

  # this is how it was done for resistance paper1
  # nScenarios <- 10  
  # ## do model runs
  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
  # ## save results objects as rda for re-use
  # save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))

>>>>

n_scenarios <- 100

@@

for(i in 1:n_scenarios)
{
  # inputs (could also be saved to a dataframe first & extracted one row at a time)
  linputs <- list()
  # constant inputs
  # for Ians Q1 set coverage to 1 & cost to 0 
  linputs$coverage <- 1
  #linputs$migration <- 0.1
  linputs$cost <- 0.1 #0
  linputs$max_gen <- 400
  # variable inputs
  linputs$expo_hi <-  runif(1, min=0.1, max=0.9)
  linputs$eff <-      runif(1, min=0.3, max=1)

  # because I want to compare sequence rot_when_r to rotation
  # I could run all the scenarios above for both seq & rot
  # (and filter out later)

  #random rot interval integer
  linputs$rot_interval <- sample( 5:50, 1 )  #beware set replace=T if more than 1

  for(rot_or_not in 0:1)
  {  
    #to get rot_int of 0 every other run
    if (rot_or_not == 0) linputs$rot_interval <- 0

    ##################
    # run one scenario
    dfres <- do.call(run_rot, linputs)

    # summarise results per scenario
    # NOTE this is also currently calc at end of run_rot() prob better to calc there
    # and put in the output dfres, but its then repeated per 
    # TODO first AG fix this maybe go for non dplyr solution
    # if I do calc in run_rot then just find value at first row for each gene
    # Error in dplyr::filter(active_or_refuge == "active") : 
    # object 'active_or_refuge' not found
    res <- dfres %>%
    # only assess in control areas not in refugia  
    dplyr::filter(active_or_refuge=='active') %>%
    group_by(resist_gene) %>%
    # for all insecticides in all generations  
    # summarise(gens_under50 = sum(resistance < 0.5, na.rm=TRUE)) %>%
    # summarise(mean_gens_under50 = mean(gens_under50)) %>%
    # just for deployed insecticides 
    summarise(gens_dep_under50 = sum(resistance < 0.5 &
                                     #finds insecticide in use = this one
                                     resist_gene==paste0('insecticide',insecticide), na.rm=TRUE)) %>%
                                 summarise(tot_gens_dep_under50 = sum(gens_dep_under50)) %>%    
                                 unlist()

    #print scenario num and inputs to doc for exploration
    cat("scenario",i,
        " expo_hi",signif(linputs$expo_hi,2),
        " eff",signif(linputs$eff,2),
        " rot_interval",linputs$rot_interval,
        "\n")

    #print results in doc for diagnostics
    cat("tot gens deployed under freq 0.5 =", res,"\n\n")

    # create dataframe from inputs and outputs
    # first add scenario column onto results
    res$scenario <- i
    df_in_out <- as.data.frame(c(linputs, res))

@@

  } #end rot_or_not loop
} #end scenario loop

# save object containing inputs and results as rda for analysis
save(df_res_all, file='df_res_all_rot.rda') #paste0(outFolder,'*.rda'))

@@

\pagebreak

# summary plot of all runs, would show any differences between sequence & rotations as non horizontal lines

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=6}

# want to plot line between sequence and rotation value
# any deviation from horizontal will be obvious
# could use rotation interval as the x axis, so that longer intervals will be clearer

# maybe not for publication
# but useful for quick view

    gg <- ggplot( df_res_all, aes_string(x='rot_interval',
                                         y='tot_gens_dep_under50',
                                         group='scenario',
                                         colour='scenario') ) +
          geom_line( alpha=0.5, lwd=1 )

    plot(gg)
